---
title: "Where Might the Mighty MadBum Move?"
author: "Evan Howard"
date: "2023-04-20"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(baseballr)
library(knitr)
library(kableExtra)
library(tidyverse)
library(pitchRx)
library(dplyr)
library(formattable)
library(reshape)
```

## MadBum's 2022 Season

With the recent news that MadBum has been DFA by the Diamondbacks, I wanted to look at what teams might be interested in picking him up. All the data that I use is from [Baseball Reference](https://www.baseball-reference.com/players/b/bumgama01.shtml) and [Baseball Savant](https://baseballsavant.mlb.com/savant-player/madison-bumgarner-518516?stats=statcast-r-pitching-mlb). I'm still very new to R and I realize that much of my code could probably be done easier. Yet, with time and practice I know that I'll get there. As such, when it comes to the table below, the column named 'n' represents Whiff%.

```{r, message=FALSE, results = 'hide'}
playerid_lookup(last_name = "Bumgarner", first_name = "Madison")
#bbref_id = bumgama01, fangraphs_id = 5524, retrosheet_id = bumgm001, mlbam_id = 518516

madbum_pitching <- statcast_search_pitchers(start_date = "2022-03-01", 
                                            end_date = "2023-04-19",
                                             pitcherid = 518516)

fg_mad <- baseballr::fg_pitcher_game_logs(playerid = 5524, year = "2023")
```

```{r, echo = FALSE}
#used to calculate whiff%
swstrike <- madbum_pitching[grep("swinging_strike", 
                                           madbum_pitching$description),]
#used to calculate whiff%
hit_into_play <- madbum_pitching[grep("hit_into_play", 
                                       madbum_pitching$description), ]

#used to calculate whiff%
foul <- madbum_pitching[grep("foul", 
                              madbum_pitching$description), ]

#used to calculate whiff%
foul_tip <- madbum_pitching[grep("foul_tip", 
                                  madbum_pitching$description), ]
#used to calculate whiff%
whiff <- data.frame(round((count(swstrike)/(count(swstrike) + count(hit_into_play) + count(foul) +
                           count(foul_tip))),3))

mb_df <- data.frame(Player = c("Madison Bumgarner"), Season = c("2023"),
                    W = c(sum(fg_mad$W)), L = c(sum(fg_mad$L)), 
                    WLperc = round(c(sum(fg_mad$W))/ (c(sum(fg_mad$W)) + c(sum(fg_mad$L))),3),
                    ERA = round(c((sum(fg_mad$ER)/sum(fg_mad$IP)*9)),2), G = c(sum(fg_mad$Games)),
                    IP = c(sum(fg_mad$IP)), H = c(sum(fg_mad$H)), R = c(sum(fg_mad$R)), 
                    ER = c(sum(fg_mad$ER)), HR = c(sum(fg_mad$HR)), BB = c(sum(fg_mad$BB)), 
                    SO = c(sum(fg_mad$SO)), Kper9 = round(c(sum(fg_mad$SO)/(sum(fg_mad$IP)/9)),1), 
                    BBper9 = round(c(sum(fg_mad$BB)/(sum(fg_mad$IP)/9)),1), 
                    Hper9 = round(c(sum(fg_mad$H))/(sum(fg_mad$IP)/9),1), 
                    KBBratio = round(c(sum(fg_mad$SO))/ c(sum(fg_mad$BB)),1),
                    WHIP = round((c(sum(fg_mad$BB)) + c(sum(fg_mad$H))) / c(sum(fg_mad$IP)),3),
                    Whiff = c(whiff))

kable(mb_df) %>%
  kable_styling("striped", full_width = TRUE) %>%
  scroll_box(width = "100%", height = "115px")

##kable(mb_df, booktabs = TRUE) %>%
  ##kable_styling(latex_options = "scale_down")
  
```

I chose to focus on two teams, one from the American League and one from the National League. The AL team I chose is the Oakland A's and the NL team is the Cincinnati Reds (5 guys on IL, Luis Cessa 5 guy in 7 for ERA, Connor Overton on IL 10 for ERA, Levi Stoudt who is 4 in ERA is in the Minors).

```{r, results = 'hide'}
ALSta <- data.frame(fg_team_pitcher(2023, 2023, league = "al", pitcher_type = "sta"))
NLSta <- data.frame(fg_team_pitcher(2023, 2023, league = "nl", pitcher_type = "sta"))
```
## Oakland A's
The Oakland A's are currently last in the AL West, and last in the AL with a 3-16 record. Their starting pitchers have gone 0-9 with a 8.59 ERA. On their 40-man roster they have 7 pitchers who are on the IL. Both [James Kaprielian](https://www.mlb.com/player/james-kaprielian-621076) and [Shintaro Fujinami](https://www.mlb.com/player/shintaro-fujinami-660261) are 0-2 and 0-3 with a 12.15 and 11.37 ERA, respectively. Currently, they rank eigth and eleventh in the league by largest ERA.
```{r, echo=FALSE}
OAK<- ALSta %>% filter(ALSta$Team == "OAK")
OAK <- OAK[,c(1,3:7, 13, 15:19,24, 29:31,36:39,42)]

colnames(OAK)[2] <- "Player"
OAK$WLperc <- (OAK$W / (OAK$W + OAK$L))
OAK <- OAK[,c(2,1,3:4, 22, 5:21)]
OAK <- OAK[,c(-16:-18)]
OAK <- OAK[,c(1:4, 6,7:19,5)]
OAK <- OAK[,c(1:16, 18, 17, 19)]
mb_df <- mb_df[,c(-20)]
names(mb_df)[15:18] <- c("K_9", "BB_9", "H_9", "K_BB")

kable(OAK) %>%
  kable_styling("striped", full_width = TRUE) %>%
  scroll_box(width = "100%", height = "90px")


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
